---
title: "Welcome to my 2022 Projections for Hitters 6x6"
author: "Darshan Patel"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_float: true
    number_sections: true
    theme: sandstone
    highlight: tango
    fig_caption: true
    df_print: paged
---

<html>

<p>

Latest Projections using XGboost

</p>

<p>

Estimated with R Studio

</p>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/Admin/Documents/Learning Python Folder1/Python Essence Training/Fantasy-Baseball')
options(knitr.table.format = "html") 
options(digits=2)
```

# Project Scope

This project is designed to showcase how Using a Rank Based Worth System values Fantasy Baseball Players

INSERT TABLE of Projections, etc, to summarize project  

***  

# Processing the Data {.tabset}

## Getting Data Into R

### Load Libraries

<p style="color:black;">

*First we need to load the packages that R needs to run the analysis*

</p>

```{r load library,message = FALSE,warning=FALSE}
library(sqldf) #SQL in R
library(skimr) #Summaries and useful for removing low % data
library(ggplot2) #Plotting Functions
library(plyr) #slightly deprecated data cleaning
library(dplyr) #slightly updated data cleaning
library(tidyverse) #tidyverse data cleaning universe
library(caret) #wrapper for creating, tuning and validating models
library(xgboost) #package for creating regression tree model
library(vtreat) # useful package for treating data before modeling 
library(Matrix)
library(Boruta)
library(mgcv)
library(moments) #for measuring skewness
library(data.table) #alternative to dplyr we use to create lags
library(pdp) #partial dependence graphs
library(vip) #variable importance 
library(grid) #put multiple plots on one grid
library(gridExtra) #additional grid functionality
library(janitor) #one function used to clean transposed data set
library(ggpubr) #for qq plot
library(tableHTML)
library
library(owmr)
```

The \# comments generally explain what additional functionality each library adds to R

### Load in Data

All data is downloaded from Fan Graphs. From this [location](https://www.fangraphs.com/leaders.aspx?pos=all&stats=bat&lg=all&qual=y&type=8&season=2021&month=0&season1=2021&ind=0). The data is also available on my Github [here](https://github.com/dissipation/Fantasy-Baseball). There are player level and team data sets

```{r data read-in, results= 'hide',message=FALSE}

#data read-in
Batter_data <- read_csv("FanGraphs Leaderboard_Hitting50PA.csv")

#Team datasets
FDG_Team = read_csv("FanGraphs Leaderboard_Team.csv")


#Create a prefix for all team stats that starts with T_
FDG_Team2 <- FDG_Team %>% 
  rename_with( ~ paste0("T_", .x))

```

### Checking Team Data

`str` give information about an object, while `skim` provides a customizable summary  

```{r checking team data}

#Output not shown for space
#str(FDG_Team2)

skim(FDG_Team2) %>%  
  tibble::as_tibble()


```
***  

## Understanding the Dataset

### Exploring the dataset

`skim` let's us see how the data was imported into R. Documentation can be found [here](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html)

```{r}

#Full Dataset dimensions

skimr::skim(Batter_data) %>% 
  tibble::as_tibble() %>% 
  select(skim_type,skim_variable,complete_rate) %>% 
  filter(complete_rate >0.30) #288 Variables

#skim_type - character or numeric
#skim_variable - name of variable
#complete_rate - % of data that is not missing
#filter - only keep variables that have 30% of data populated
```

Additionally let's look at how variables vary by year to see if there are any discrepancies there  

```{r}

#It looks like one year, there were fewer games played, and there is a clear drop off in home runs
Batter_data_dist =
Batter_data %>% 
 group_by(Season) %>% 
  summarize (Games_played = max(G),
             Avg_HR= mean(HR)
             )
Batter_data_dist

ggplot(Batter_data_dist, aes(Season, Avg_HR)) +
  geom_col()+
  ggtitle("Average Home Runs by Year")+
  theme(plot.title = element_text(hjust = 0.5,size = 22,color ="steel blue"))



```

***  

## Cleaning and Creating Initial Dataset for Model

What are some issues with the data?

1.  Many of Variables, such as K%, are being read in as characters

    -   Only Team and Player Name should be characters

2.  There is spotty data coverage in some of the variables (\~Variables have less than 30% Coverage)

3.  2020 Data only includes 60 games worth of data

    -   This was a season shortened due to Covid-19

4.  Team Data needs to be appended to Batter Data by Team Name 

***  

### **Cleanly Changing all Variables that are characters to numeric.**
There are several ways to do this, we will identify the variables we want to change that are mis-identified. `parse_number` can be used to pull numbers from these variables. Additional ways to tackle this can be found [here](https://stackoverflow.com/questions/8329059/how-to-convert-character-of-percentage-into-numeric-in-r)

```{r}

#Select Column names that are characters but not Team or Name, These should be percentages
Batter_data_chars_to_convert <- Batter_data %>% 
  select_if(is.character)%>% select(-Team,-Name) %>% 
  mutate_all (function(x) as.numeric(readr::parse_number(x))/100)
#Note : There are additional ways to do this, this is just one solution


#We can exclude the variables we converted and reintroduce them
Batter_data_num <- Batter_data %>% select(-colnames(Batter_data_chars_to_convert))

Batter_data2 = cbind(Batter_data_num,Batter_data_chars_to_convert) %>% 
  select (colnames(Batter_data)) %>%  #preserve original order 
  dplyr::rename(flyball_perc = `FB%...46`,fastball_perc = `FB%...73`) #rename two ambiguous columns
  
skim(Batter_data2) %>% 
  as_tibble() %>% 
  group_by(skim_type) %>% 
  count()


#Logical variables are R's best guess, in our case they are all NA's and will be removed

```

The same can be done for the Team Data that is loaded  


```{r}

#Select Column names that are characters but not Team or Name, These should be percentages
FDG_Team2_chars_to_convert <- FDG_Team2 %>% 
  select_if(is.character)%>% select(-T_Team) %>% 
  mutate_all (function(x) as.numeric(readr::parse_number(x))/100)
#Keep in mind, parse number may make actual characters into numerical variables so carefully check your data before using

#We can exclude the variables we converted and reintroduce them
FDG_Team2_num <- FDG_Team2 %>% select(-colnames(FDG_Team2_chars_to_convert))

FDG_Team3 = cbind(FDG_Team2_num,FDG_Team2_chars_to_convert) %>% 
  select (colnames(FDG_Team2)) %>%  #preserve original order
dplyr::rename(T_flyball_perc = `T_FB%...45`,T_fastball_perc = `T_FB%...72`) 

skim(FDG_Team3) %>% 
  as_tibble() %>% 
  group_by(skim_type) %>% 
  count()



```
***  


### Filtering Data with Low Coverage    
I choose 30% coverage of data necessary but this can be adjusted up or down. This will also get rid of columns that are all `NA`.  
```{r}

# Keep variables with enough values (Need 30% data coverage rate here)
Player_cols_to_keep =
skim(Batter_data2) %>% 
  dplyr::select(skim_type, skim_variable, complete_rate) %>% 
  filter (complete_rate > 0.30)

#Transpose Rows to get column names as skim melts the data
Player_cols_to_keep_transpose = t(Player_cols_to_keep) 

#extract the colnames we would like to keep
Player_cols_to_keep = colnames(janitor::row_to_names(Player_cols_to_keep_transpose,row_number = 2))

#Only keep the columns designated to have over 30% of their data populated or greater
Batter_data3 = Batter_data2 %>% 
  select(one_of(Player_cols_to_keep)) 


```


*Repeat the process for Team Variables*
```{r}
Team_cols_to_keep =
skim(FDG_Team3) %>% 
  dplyr::select(skim_type, skim_variable, complete_rate) %>% 
  filter (complete_rate > 0.30)


#Transpose Rows to get column names as skim melts the data
Team_cols_to_keep_transpose = t(Team_cols_to_keep) 

#extract the colnames we would like to keep
Team_cols_to_keep = colnames(janitor::row_to_names(Team_cols_to_keep_transpose,row_number = 2))

#Only keep the columns designated to have over 30% of their data populated or greater
FDG_Team4 = FDG_Team3 %>% 
  select(one_of(Team_cols_to_keep)) 



```



***  

### Creating Variables Normalized by Year  
Some Variables will need to be normalized by Plate Appearances (PA) if they aren't a percentage already. Remaining Variables are percentages or indicies so will not need to be transformed

```{r}

Batter_data4 = Batter_data3 %>% 
  mutate( #create new variables based on existing variables
    H_PA = H/PA,
    x1B_PA = `1B`/PA, #note: R can't have variables start with a number
    x2b_PA = `2B`/PA,
    x3b_PA = `3B`/PA,
    HR_PA = HR/PA,
    R_PA = R/PA,
    RBI_PA = RBI/PA,
    BB_PA = BB/PA,
    IBB_PA = IBB/PA,
    SO_PA=SO/PA,
    HBP_PA=HBP/PA,
    SF_PA=SF/PA,
    SH_PA=SH/PA,
    GDP_PA= GDP/PA,#ground into double play
    SB_PA=SB/PA,
    CS_PA=CS/PA,
    GB_PA = GB/PA,   #Groundballs
    FB_PA =  FB/PA,  #FlyBalls
    LD_PA = LD/PA,   #LineDrives
    IFFB_PA = IFFB/PA,  #Infield Fly balls
    Pitches_PA= Pitches/PA,
    Balls_PA= Balls/PA,
    Strikes_PA= Strikes/PA,
    IFH_PA= IFH/PA,
    BU_PA= BU/PA,
    BUH_PA= BUH/PA,
    PH_PA= PH/PA,
    Barrels_PA= Barrels/PA,
    HardHits_PA= HardHit/PA
  ) %>% select(-(H:CS),-(GB:BUH),-PH,-Barrels,-HardHit,-Events) #Drop the old variables

#skim(Batter_data4) %>% as_tibble()


```

*Repeat the process for Team Variables*
```{r}

FDG_Team5 = FDG_Team4 %>% 
  mutate( #create new variables based on existing variables
    T_H_T_PA = T_H/T_PA,
    T_x1B_T_PA = T_1B/T_PA, #note: R can't have variables start with a number
    T_x2b_T_PA = T_2B/T_PA,
    T_x3b_T_PA = T_3B/T_PA,
    T_HR_T_PA = T_HR/T_PA,
    T_R_T_PA = T_R/T_PA,
    T_RBI_T_PA = T_RBI/T_PA,
    T_BB_T_PA = T_BB/T_PA,
    T_IBB_T_PA = T_IBB/T_PA,
    T_SO_T_PA=T_SO/T_PA,
    T_HBP_T_PA=T_HBP/T_PA,
    T_SF_T_PA=T_SF/T_PA,
    T_SH_T_PA=T_SH/T_PA,
    T_GDP_T_PA= T_GDP/T_PA,#ground into double play
    T_SB_T_PA=T_SB/T_PA,
    T_CS_T_PA=T_CS/T_PA,
    T_GB_T_PA = T_GB/T_PA,   #Groundballs
    T_FB_T_PA =  T_FB/T_PA,  #FlyBalls
    T_LD_T_PA = T_LD/T_PA,   #LineDrives
    T_IFFB_T_PA = T_IFFB/T_PA,  #Infield Fly balls
    T_Pitches_T_PA= T_Pitches/T_PA,
    T_Balls_T_PA= T_Balls/T_PA,
    T_Strikes_T_PA= T_Strikes/T_PA,
    T_IFH_T_PA= T_IFH/T_PA,
    T_BU_T_PA= T_BU/T_PA,
    T_BUH_T_PA= T_BUH/T_PA,
    T_PH_T_PA= T_PH/T_PA,
    T_Barrels_T_PA= T_Barrels/T_PA,
    T_HardHits_T_PA= T_HardHit/T_PA
  ) %>% select(-(T_H:T_CS),-(T_GB:T_BUH),-T_PH,-T_Barrels,-T_HardHit,-T_Events) #Drop the old variables


#skim(FDG_Team5) %>% as_tibble()


```

***  

### Creating Lagged Variables  
There are several ways to lag a dataset **BY GROUP**.    
* `Dplyr` way is [here.](https://statisticsglobe.com/create-lagged-variable-by-group-in-r).   
* While data.table (the method used below) is [here.](https://stackoverflow.com/questions/26291988/how-to-create-a-lag-variable-within-each-group)  

```{r}
#Note we will only be lagging the player level data, as the previous year's team performance shouldn't impact current performance


#Order the dataset by lag columns
Batter_data5 =  arrange(Batter_data4, playerid,Season) #playerid is the Fangraph id assigned to each player

# Convert dataframe to data.table format
DT_batter = data.table(Batter_data5)

#designate columns to lag - which is all of them
cols1 = colnames(Batter_data5)
anscols = paste("lag", cols1, sep="_")
DT_batter[, (anscols) := data.table::shift(.SD, 1, NA, "lag"),by ='playerid', .SDcols=cols1] #Create 1 period lags by year

Batter_data6 = as.data.frame(DT_batter) %>% select(-lag_playerid, -lag_Team, -lag_Season, -lag_Age,-lag_Name)

ncol(Batter_data5) #287 - no lags
ncol(Batter_data6) #574 - lagged data ~ (287 * 2)-5

```

***    

### Merging Team and Player Data  
We can use either the `merge` function or the SQL functionality provided by the `sqldf` package to join the lagged player level data to the Team level data

```{r}

df_batting_init = sqldf(
  "
  select a.*, b.*
  from Batter_data6 a
  left join FDG_Team5 b
  on a.Team = b.T_Team and a.Season = b.T_Season
  
  "
)  %>% select(-T_Team,-T_Season,T_Age,T_G,T_AB)# Unncessary Team Variables


nrow(df_batting_init) - nrow(Batter_data6) #check if any rows are duplicated


```


***  


# Creating Rankings for Players Based On Percentiles {.tabset}  
We can use Percentile based ranking to get rankings for players from the 2021 season.  

## Worth of each stat  

### Calculating past performance   

Each player goes from a 0% to 100% on each percentile stat that is used for creating a scoring opportunity. All data is already normalized by plate appearances, but must now be ranked for each year.   
*Code here is for a 6x6 league I am in that includes OPS, but projections on main page are based on a standard 5x5 league*  

```{r}

#Categories I include are:
#Runs (R), Home Runs (HR), Runs Batted In (RBI), Stolen Bases (SB), Batting Average (AVG), On-base + Slugging Percentage (OPS)

df_batting_init2 =  df_batting_init %>%
#  arrange(player_id,year) %>% 
  group_by(Season) %>% 
  mutate(
    Runs_share = order(order(rank(R_PA,ties.method = 'average'),decreasing = FALSE))/n(),
     HR_share = order(order(rank(HR_PA,ties.method = 'average'),decreasing = FALSE))/n(),
     RBI_share = order(order(rank(RBI_PA,ties.method = 'average'),decreasing = FALSE))/n(),
     SB_share = order(order(rank(SB_PA,ties.method = 'average'),decreasing = FALSE))/n(),
     AVG_share = order(order(rank(AVG,ties.method = 'average'),decreasing = FALSE))/n(),
    OPS_share = order(order(rank(OPS,ties.method = 'average'),decreasing = FALSE))/n(),
    Worth = Runs_share+HR_share+RBI_share+SB_share+AVG_share+OPS_share
    ) %>% 
  ungroup() %>% 
select(-R_PA,-HR_PA,-RBI_PA,-SB_PA,-AVG,-OPS)

```

Chart of the Distribution of initial percentiles  
As the chart below shows, the data is roughly normal.
```{r}

skewness((df_batting_init2$Worth))

ggplot2::qplot(df_batting_init2$Worth, main="Total Dataset") + geom_histogram(colour="black", fill="grey") + theme_bw()

min(df_batting_init2$Worth)

max(df_batting_init2$Worth)

ggpubr::ggqqplot(df_batting_init2$Worth)

shapiro.test(df_batting_init2$Worth)
```


***  


## 2021 Player Rankings per Plate Appearance 
### 2021 Player Rankings - Top Worth Players with OPS
Total Rankings for the players (Using 5x5 Scoring) can be found [here]()  
```{r,warning=FALSE}

#Categories I include are:
#Runs (R), Home Runs (HR), Runs Batted In (RBI), Stolen Bases (SB), Batting Average (AVG), On-base + Slugging Percentage (OPS)

df_batting_init2 =  df_batting_init %>%
#  arrange(player_id,year) %>% 
  group_by(Season) %>% 
  mutate(
    Runs_share = order(order(rank(R_PA,ties.method = 'average'),decreasing = FALSE))/n(),
     HR_share = order(order(rank(HR_PA,ties.method = 'average'),decreasing = FALSE))/n(),
     RBI_share = order(order(rank(RBI_PA,ties.method = 'average'),decreasing = FALSE))/n(),
     SB_share = order(order(rank(SB_PA,ties.method = 'average'),decreasing = FALSE))/n(),
     AVG_share = order(order(rank(AVG,ties.method = 'average'),decreasing = FALSE))/n(),
    OPS_share = order(order(rank(OPS,ties.method = 'average'),decreasing = FALSE))/n(),
    Worth = Runs_share+HR_share+RBI_share+SB_share+AVG_share+OPS_share
    ) %>% 
  ungroup() %>% 
select(-R_PA,-HR_PA,-RBI_PA,-SB_PA,-AVG,-OPS)



options(digits=2)

df_batting_init2021 =
df_batting_init2 %>% 
  group_by(Name) %>% 
  filter(Season == 2021) %>% 
  arrange(desc(Worth)) %>% 
  select(Name,Runs_share,HR_share,RBI_share, SB_share,OPS_share,AVG_share,Worth)


df_batting_init2021 %>%
  filter (Worth>3.9) %>% 
  kbl() %>% 
 kable_material(c("striped", "hover","condensed","responsive"),full_width = F,fixed_thead = T)

```


***  

## 2021 Player Rankings Total EoY Ranks   
### 2021 Player Rankings - Top Worth Players with OPS   
Total Rankings for the players (Using 5x5 Scoring) can be found [here]()  
```{r,warning=FALSE}


#Categories I include are:
#Runs (R), Home Runs (HR), Runs Batted In (RBI), Stolen Bases (SB), Batting Average (AVG), On-base + Slugging Percentage (OPS)

df_batting_init_raw =  Batter_data3 %>%
#  arrange(player_id,year) %>% 
  group_by(Season) %>% 
  mutate(
    Runs_share = order(order(rank(R,ties.method = 'average'),decreasing = FALSE))/n(),
     HR_share = order(order(rank(HR,ties.method = 'average'),decreasing = FALSE))/n(),
     RBI_share = order(order(rank(RBI,ties.method = 'average'),decreasing = FALSE))/n(),
     SB_share = order(order(rank(SB,ties.method = 'average'),decreasing = FALSE))/n(),
     AVG_share = order(order(rank(AVG,ties.method = 'average'),decreasing = FALSE))/n(),
    OPS_share = order(order(rank(OPS,ties.method = 'average'),decreasing = FALSE))/n(),
    Worth = Runs_share+HR_share+RBI_share+SB_share+AVG_share+OPS_share
    ) %>% 
  ungroup() %>% 
select(-R,-HR,-RBI,-SB,-AVG,-OPS)



options(digits=2)

df_batting_init2021_raw =
df_batting_init_raw %>% 
  group_by(Name) %>% 
  filter(Season == 2021) %>% 
  arrange(desc(Worth)) %>% 
  select(Name,Runs_share,HR_share,RBI_share, SB_share,OPS_share,AVG_share,Worth)


df_batting_init2021_raw %>%
  filter (Worth>3.9) %>% 
  kbl() %>% 
 kable_material(c("striped", "hover","condensed","responsive"),full_width = F,fixed_thead = T)


```




# Creating Model File   
## Additional Data Prep  
### Remove Variables which are based off current hitting numbers    
Not all variables can be used for predictive modeling.   
```{r}
df_batting_init3 = df_batting_init2 %>% 
  select(-`HR/FB`, -Dol,-`Age Rng`,PA,-(H_PA:PH_PA),-Name)
```

Lag Share Variables to use for predictive modeling. The variables that we created for the Worth metric must also be removed. This will create the final dataset.

```{r}

#Order the dataset by lag columns
df_batting_init4 =  arrange(df_batting_init3, playerid,Season) #playerid is the Fangraph id assigned to each player

# Convert dataframe to data.table format
DT_batter2 = data.table(df_batting_init4)

#designate columns to lag - which is all of them
cols1 = (c('Runs_share','HR_share','RBI_share', 'SB_share','OPS_share','AVG_share','Worth'))
anscols = paste("lag", cols1, sep="_")
DT_batter2[, (anscols) := data.table::shift(.SD, 1, NA, "lag"),by ='playerid', .SDcols=cols1] #Create 1 period lags by year

df_batting_final = as.data.frame(DT_batter2) %>% 
  select(-c(Runs_share,HR_share,RBI_share, SB_share,OPS_share,AVG_share))

```



### Creating Training/Test Split  
We split the data into Training Data (which is used to create the model) and test data (which is used to validate the model)   
```{r}

set.seed(15674)  # For reproducibility
# Create index for testing and training data
inTrain <- createDataPartition(y = df_batting_final$Worth, p = 0.80, list = FALSE)
# subset pitching data for training
tr_2021 <- df_batting_final[inTrain,]
# subset the rest to test and validate trained model
te_2021 <- df_batting_final[-inTrain,]

nrow(tr_2021)/nrow(df_batting_final) #check if split is 0.8

```

### Treat Missing Data by Imputing Mean Value  
Vtreat Package in R is excellent for treating data before using for modeling. Additional documentation can be found [here.](https://winvector.github.io/vtreat/index.html)
```{r}
treat_plan_2021 <- vtreat::designTreatmentsZ(
  dframe = tr_2021, # training data
  varlist = colnames(tr_2021) %>% .[. != "hitting_score1"], # input variables = all training data columns, except random
  codeRestriction = c("clean", "isBAD", "lev"), # derived variables types (drop cat_P)
  verbose = FALSE) # suppress messages

#clean stands for cleaned numerical variable, isBAD indicates that a value replacement has occurred (which indicates a missing value in this case), and lev is a binary indicator whether a particular value of that categorical variable was present.  

#### Checking Scoreframe

score_frame <- treat_plan_2021$scoreFrame %>% 
  select(varName, origName, code)

head(score_frame)


tr_treated_2021 <- vtreat::prepare(treat_plan_2021, tr_2021)
te_treated_2021 <- vtreat::prepare(treat_plan_2021, te_2021)
total_treated_2021_hitting <- vtreat::prepare(treat_plan_2021, df_batting_final)

#tr_treated = tr
#te_treated = te

dim(tr_treated_2021) #note there are dummies for each player and team

```


***    


### Check Distribution of Training Population  
The population used for Training should be indicative of Total Population
```{r}

ggplot2::qplot(tr_treated_2021$Worth, main="Training Set") + geom_histogram(colour="black", fill="grey") + theme_bw()

skewness(tr_treated_2021$Worth) #The skewness is the same as the overall


```


# Running XGboost Model {.tabset} 
To keep things simple with modeling, we’ll turn the training data into simple input variables for `caret::train`, dropping the response variable and converting the data frame to a matrix. Documentation for this approach to XGboost can be found [here.](https://www.kaggle.com/pelkoja/visual-xgboost-tuning-with-caret)    

## Tuning the Model

### Initial Non-Tuned Model
Break the data set into x and y inputs with x being a matrix  
```{r}
input_x <- as.matrix(((tr_treated_2021))%>%
   select(-Worth) %>%                      
   select(!ends_with ("_isBAD")))

input_y <- tr_treated_2021$Worth

```

XGBoost with Default Hyperparameters    
The Variable Importance (`caret::varImp(xgb_base_2021, scale = F  )`) from the caret package shows the contribution of each variable to the initial model. As you can see SLG_plus_ (SLG+) takes up much of the importance as it is derived from SLG (one of the key contributors to Worth). These types of variables will be removed during variable selection in the next step.  
*XGBoost documentation can be found for more general models [here.](https://www.kaggle.com/code/rtatman/machine-learning-with-xgboost-in-r/notebook)*

```{r}

#Defaults for xgboost model
grid_default <- expand.grid(
  nrounds = 100,
  max_depth = 6,
  eta = 0.3,
  gamma = 0,
  colsample_bytree = 1,
  min_child_weight = 1,
  subsample = 1
)

#This is a blank train_control set, this will be updated after
train_control <- caret::trainControl(
  method = "none",
  verboseIter = FALSE, # no training log
  allowParallel = TRUE # FALSE for reproducible results 
)

xgb_base_2021 <- caret::train(
  x = input_x,
  y = input_y,
  trControl = train_control,
  tuneGrid = grid_default,
  method = "xgbTree",
  verbose = TRUE
)

caret::varImp(xgb_base_2021, scale = F  )




```

## Further Variable Selection  
### Remove redundant and highly correlated variables  


Selection Removal Step 1: Check for high correlations  
Normally, this step is done early, but those steps were reserved for preparing the data  

```{r}

dep_cor1 <- t(as.data.frame(cor(tr_treated_2021[ , colnames(tr_treated_2021) != "Worth"],
                tr_treated_2021$Worth)))
dep_cor1 <-
as.data.frame(t(as.data.frame(dep_cor1)%>% 
  select(!starts_with("lag")) %>% #remove lag variables
  select(!contains("_isBAD")))) 

dep_cor1 <- tibble::rownames_to_column(dep_cor1,"VARIABLES")%>% #remove indicators for missing data
  filter(V1 > 0.70|V1 < -0.5)

dep_cor1

dep_cor2 <- colnames(row_to_names(t(dep_cor1),row_number = 1))



```
Let's Remove variables with high correlation to worth metric, and metrics that are calculated after a player's performance (such as WAR)

```{r}

input_x <- as.matrix(((tr_treated_2021))%>%
   select(-Worth) %>% #Remove dependant variables
     select (-all_of(dep_cor2),-RAR,-WAR,-WPA,-Spd,-Rep,-BABIP,-Def,-BABIP_plus_,-PA,-AB,-G,-wFB,	-wSL,	-wCT,	-wCH,	-wSF, -X_plus_WPA, -wSB, -BsR,-Pos,-UBR , -wCB #Remove redundant variables or non/weighted variables
) %>%      
select(!ends_with ("_isBAD"))) #indicator variable for missing data

input_y <- tr_treated_2021$Worth





```

Run the model on the new dataset to make sure the variable importances look fine
```{r}

#Note Training parameters were set in initial model set up
xgb_base_2021 <- caret::train(
  x = input_x,
  y = input_y,
  trControl = train_control,
  tuneGrid = grid_default,
  method = "xgbTree",
  verbose = TRUE
)

caret::varImp(xgb_base_2021, scale = F  )


```


## Model with new data  

### Tuning All Hyperparameters
A tune grid allows us to test a large amount of hyper-parameters and find the model with the lowest RMSE for predictions.   
However, The more values you want to test and the greater the amount of Cross-Fold Validations (`method = "cv"`), the greater the computational time it will take. More information on the specific parameters can be found [here.](https://www.hackerearth.com/practice/machine-learning/machine-learning-algorithms/beginners-tutorial-on-xgboost-parameter-tuning-r/tutorial/)

```{r}

# maximum number of trees
nrounds <- 1000

# note to start nrounds from 200, as smaller learning rates result in errors so
# big with lower starting points that they'll mess the scales
tune_grid <- expand.grid(
  nrounds = seq(from = 100, to = nrounds, by = 50),
  eta = c(0.01, 0.025, 0.05, 0.1),
  max_depth = c(2, 4, 6, 8),
  gamma = 0,
  colsample_bytree = 1,
  min_child_weight = 1,
  subsample = 1
)

tune_control <- caret::trainControl(
  method = "cv", # cross-validation
  number = 5, # with n folds 
  ## Note this was # out in the original code
  #index = createFolds(tr_treated$Id_clean), # fix the folds
  verboseIter = FALSE, # no training log
  allowParallel = TRUE # FALSE for reproducible results 
)



```

*Running the initial tuning model*  
```{r}
#Note I will be timing these runs to give an estimate on how long this model takes to run
start_time <- Sys.time()

xgb_tune_2021 <- caret::train(
  x = input_x,
  y = input_y,
  trControl = tune_control,
  tuneGrid = tune_grid,
  method = "xgbTree",
  verbose = FALSE
  ,verbosity = 0
)

end_time <- Sys.time()

end_time - start_time

```

*Tuning Plot and Variable Importance*
```{r}
varImp(xgb_tune_2021, scale = F  ) 


# helper function for the plots
tuneplot <- function(x, probs = .90) {
  ggplot(x) +
    coord_cartesian(ylim = c(quantile(x$results$RMSE, probs = probs), min(x$results$RMSE))) +
    theme_bw()
}

tuneplot(xgb_tune_2021)
```

### Fine Tuning Model  
#### Second Tuning: Maximum Depth and Minimum Child Weight  
After fixing the learning rate to 0.1 and we’ll also set maximum depth to 3 +-1 (or +2 if max_depth == 2) to experiment a bit around the suggested best tune in previous step. Then, well fix maximum depth and minimum child weigh

```{r}
tune_grid2 <- expand.grid(
  nrounds = seq(from = 50, to = nrounds, by = 50),
  eta = xgb_tune_2021$bestTune$eta,
  max_depth = ifelse(xgb_tune_2021$bestTune$max_depth == 2,
    c(xgb_tune_2021$bestTune$max_depth:4),
    xgb_tune_2021$bestTune$max_depth - 1:xgb_tune_2021$bestTune$max_depth + 1),
  gamma = 0,
  colsample_bytree = 1,
  min_child_weight = c(1, 2, 3),
  subsample = 1
)

xgb_tune2_2021 <- caret::train(
  x = input_x,
  y = input_y,
  trControl = tune_control,
  tuneGrid = tune_grid2,
  method = "xgbTree",
  verbose = TRUE
)

tuneplot(xgb_tune2_2021)

xgb_tune2_2021$bestTune

varImp(xgb_tune2_2021, scale = F  ) 
```

#### Third Tuning: Column and Row Sampling

```{r}

tune_grid3 <- expand.grid(
  nrounds = seq(from = 50, to = nrounds, by = 50),
  eta = xgb_tune_2021$bestTune$eta,
  max_depth = xgb_tune2_2021$bestTune$max_depth,
  gamma = 0,
  colsample_bytree = c(0.4, 0.6, 0.8, 1.0),
  min_child_weight = xgb_tune2_2021$bestTune$min_child_weight,
  subsample = c(0.5, 0.75, 1.0)
)

xgb_tune3_2021 <- caret::train(
  x = input_x,
  y = input_y,
  trControl = tune_control,
  tuneGrid = tune_grid3,
  method = "xgbTree",
  verbose = TRUE
)

tuneplot(xgb_tune3_2021, probs = .95)

xgb_tune3_2021$bestTune

varImp(xgb_tune3_2021, scale = F  ) 

```

#### Fourth Tuning: Gamma  
Next, we again pick the best values from previous step, and now will see whether changing the gamma has any effect on the model fit:
```{r}
tune_grid4 <- expand.grid(
  nrounds = seq(from = 50, to = nrounds, by = 50),
  eta = xgb_tune_2021$bestTune$eta,
  max_depth = xgb_tune2_2021$bestTune$max_depth,
  gamma = c(0, 0.05,0.1, 0.2,0.4, 0.5, 0.7, 0.9, 1.0),
  colsample_bytree = xgb_tune3_2021$bestTune$colsample_bytree,
  min_child_weight = xgb_tune2_2021$bestTune$min_child_weight,
  subsample = xgb_tune3_2021$bestTune$subsample
)

xgb_tune4_2021 <- caret::train(
  x = input_x,
  y = input_y,
  trControl = tune_control,
  tuneGrid = tune_grid4,
  method = "xgbTree",
  verbose = TRUE
)

tuneplot(xgb_tune4_2021)

xgb_tune4_2021$bestTune

varImp(xgb_tune4_2021, scale = F  ) 
```

#### Fifth Tuning: Reducing the Learning Rate  
Now, we have tuned the hyperparameters and can start reducing the learning rate to get to the final model:  

```{r}
start_time <- Sys.time()

tune_grid5 <- expand.grid(
  nrounds = seq(from = 100, to = 10000, by = 75),
   eta = c(0.01, 0.015, 0.025,0.035, 0.05,0.75, 0.1),
  max_depth = xgb_tune2_2021$bestTune$max_depth,
  gamma = xgb_tune4_2021$bestTune$gamma,
  colsample_bytree = xgb_tune3_2021$bestTune$colsample_bytree,
  min_child_weight = xgb_tune2_2021$bestTune$min_child_weight,
  subsample = xgb_tune3_2021$bestTune$subsample
)



xgb_tune5_2021 <- caret::train(
  x = input_x,
  y = input_y,
  trControl = tune_control,
  tuneGrid = tune_grid5,
  method = "xgbTree",
  verbose = TRUE
)

#tuneplot(xgb_tune5_2021)

end_time <- Sys.time()

end_time - start_time

xgb_tune5_2021$bestTune

varImp(xgb_tune5_2021, scale = F  ) 
```


#### Fitting Final Model

```{r}

(final_grid_2021 <- expand.grid(
  nrounds = xgb_tune5_2021$bestTune$nrounds,
  eta = xgb_tune5_2021$bestTune$eta,
  max_depth = xgb_tune5_2021$bestTune$max_depth,
  gamma = xgb_tune5_2021$bestTune$gamma,
  colsample_bytree = xgb_tune5_2021$bestTune$colsample_bytree,
  min_child_weight = xgb_tune5_2021$bestTune$min_child_weight,
  subsample = xgb_tune5_2021$bestTune$subsample
))

(xgb_model_2021 <- caret::train(
  x = input_x,
  y = input_y,
  trControl = train_control,
  tuneGrid = final_grid_2021,
  method = "xgbTree",
  verbose = TRUE
))

varImp(xgb_model_2021, scale = F  ) 

```

### Model Performance  


#### Checking Model on Test Split Data  

```{r}


y_pred_test <- predict(xgb_model_2021, data.matrix(te_treated_2021))

test_stats= cbind((te_treated_2021$Worth),y_pred_test)

test_statsR2 = cor(test_stats[,1],test_stats[,2])^2

print(test_statsR2)


y_pred_train <- predict(xgb_model_2021, data.matrix(tr_treated_2021))

train_stats = cbind((tr_treated_2021$Worth),y_pred_train)

train_statsR2 = cor(train_stats[,1],train_stats[,2])^2

print(train_statsR2)

#test dataset
x <- select(te_treated_2021, -Worth)
y <- (te_treated_2021$Worth)

(xgb_model_rmse <- ModelMetrics::rmse(y, predict(xgb_model_2021, newdata = x)))

holdout_x <- select(tr_treated_2021, -Worth)
holdout_y <- tr_treated_2021$Worth

(xgb_model_rmse <- ModelMetrics::rmse(holdout_y, predict(xgb_model_2021, newdata = holdout_x)))


```

#### Graphical Representation of Model   


```{r}

ggplot2::ggplot() +
  aes(x = test_stats[,1], y = test_stats[,2]) +
  geom_jitter() +
  xlab("Predicted Values") +
  ylab("Actual Values") +
  ggtitle("Results of Pitching Model on Test Data")+
  theme(plot.title = element_text(hjust = 0.5,size = 22,color ="steel blue"))+
  geom_smooth(method = "lm")



```


# Creating 2022 Projections from Model  {.tabset} 


## Re-fit model for Important Variables
Now that we have an acceptable model, we can use it to create projections for how well we think players should do in 2022 based on their hitting statistics in 2021. First let's reduce

1. Only keep variables with high enough importance in model  

```{r}


vip(xgb_model_2021, num_features = 30)  # 10 is the default, 30 gives a visual on the top 30 most important features of the model

unscalevi = vi(xgb_model_2021, method="model") #shows the numbers behind the plot

unscalevi$Importance_perc = with(unscalevi,Importance/sum(Importance)) #adds percentages 

unscalevi # importance by variables

variables_to_keep_2021 = subset(unscalevi, Importance_perc > 0.0010) %>% select(Variable) #Keep Variables that explain at least a small amount [0.1%] of the model. This is a low threshold for inclusion ,but you can adjust this

variables_to_keep_2021b = t(variables_to_keep_2021)

variables_to_keep_2022 = colnames(row_to_names(variables_to_keep_2021b,row_number = 1))

tr_treated_2022 = tr_treated_2021 %>%  select(Worth,one_of(variables_to_keep_2022),starts_with("Team_lev_x_")) #keep modeled important variables along with team indicator variables

te_treated_2022 = te_treated_2021 %>%  select(Worth,one_of(variables_to_keep_2022),starts_with("Team_lev_x_"))

input_x_2022 = as.matrix(select(tr_treated_2022, -Worth))

input_y_2022 = tr_treated_2022$Worth



```

2. Re-fit model with reduced variable scope  

```{r}


(final_grid_2021 <- expand.grid(
  nrounds = xgb_tune5_2021$bestTune$nrounds,
  eta = xgb_tune5_2021$bestTune$eta,
  max_depth = xgb_tune5_2021$bestTune$max_depth,
  gamma = xgb_tune5_2021$bestTune$gamma,
  colsample_bytree = xgb_tune5_2021$bestTune$colsample_bytree,
  min_child_weight = xgb_tune5_2021$bestTune$min_child_weight,
  subsample = xgb_tune5_2021$bestTune$subsample
))

(xgb_model_2022 <- caret::train(
  x = input_x_2022,
  y = input_y_2022,
  trControl = train_control,
  tuneGrid = final_grid_2021,
  method = "xgbTree",
  verbose = TRUE
))


vip(xgb_model_2022, num_features = 30)

unscalevi24 = vi(xgb_model_2022, method="model")

unscalevi24$Importance_perc = with(unscalevi24,Importance/sum(Importance)) 

unscalevi24


#For anything above breaking_IP we need to create projection table by age or age bucket

#write_csv(unscalevi24,"unscalevi24.csv")

```
3. Save the model
```{r}

save(xgb_model_2022,file = '2022_Hitting6x6_Model.Rdata')

```



## Get 2022 list of players

Data for Fantasy Pros Ranking can be found[here.](https://www.fangraphs.com/projections.aspx?pos=all&stats=pit&type=atc&team=0&lg=all&players=0)  
*Data downloaded on 3-25-2022*
```{r}
#Note WAR Rankings are the Rankings, from ATC

Latest_2022_playerdata_FP = read_csv("FanGraph_Fantasy_Baseball_Hitting.csv")

head(Latest_2022_playerdata_FP)


```










</html>
