---
title: "2022 Hitting 6x6 Projections"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    theme: cerulean
    highlight: tango
    fig_caption: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = 'C:/Users/Admin/Documents/Learning Python Folder1/Python Essence Training/Fantasy-Baseball/Data')
options(knitr.table.format = "html") 
options(digits=2)
options(scipen = 100)
```


#### [Go back to the main page](https://dissipation.github.io/DarshDSPortfolio/)


All data is from [FanGraphs.](https://www.fangraphs.com/) I have no affiliation with FanGraphs, but please consider contributing to their [website](https://plus.fangraphs.com/shop/) if you found this project informative.    

Positional Scarcity is determined by [ATC WAR Projections](https://www.fangraphs.com/projections.aspx?pos=all&stats=bat&type=atc&team=0&lg=all&players=0) from 3-31-2022.  

***    

# **Pitching with Holds included** {.tabset .tabset-pills}   
Players that have an `NA` for `Model_Score` did not play in the MLB in 2021. The model had a R-Squared of 0.78. Includes the Holds Category.
```{r}
### Fill in code
```


***    

## Pitchers by Position {.tabset .tabset-pills}  
My non-normalized model scores for a player are `Model_Score`.         
 
### All P  

```{r, warning=FALSE}


FanGraph_Positions_Pitcher_Total2 =
pitching_Data_NonAdj_Projections_Total %>% 
  mutate(
    Avg_IP = mean(IP),
    AdjPredict_Score_raw = ifelse(is.na(Predict_Score),NA,Predict_Score*(IP/Avg_IP)*1)
  ) %>% select(-Avg_IP)


FanGraph_Positions_Pitcher_All_POS2 =
FanGraph_Positions_Pitcher_Total2 %>% 
  mutate(
    max_predscore= max(AdjPredict_Score_raw,na.rm = T),
    AdjPredict_Score = ifelse (is.na(AdjPredict_Score_raw),NA,AdjPredict_Score_raw),
    WAR_rank = order(order(rank(WAR,ties.method = 'average'),decreasing = TRUE)),
    AdjPredict_Score_Rank = order(order(rank(AdjPredict_Score,ties.method = 'average'),decreasing = TRUE))-sum(is.na(AdjPredict_Score)),
        Ranks_Above_ADP = ADP - AdjPredict_Score_Rank
  ) %>%
filter(IP > 19) %>%   
  select (Name,Team,Position,ADP,WAR, WAR_rank,AdjPredict_Score ,AdjPredict_Score_Rank,Ranks_Above_ADP, Positional_Scarcity)

ft_dt <- FanGraph_Positions_Pitcher_All_POS2[1:nrow(FanGraph_Positions_Pitcher_All_POS2), 1:ncol(FanGraph_Positions_Pitcher_All_POS2)] %>% 
#%>% 
#filter(AdjPredict_Score_Rank>0)
 arrange((ADP))

ft_dt$ADP <- color_tile("white", "red")(ft_dt$ADP)

ft_dt$WAR <- color_bar("lightblue")(ft_dt$WAR)


ft_dt$AdjPredict_Score<-  round((ft_dt$AdjPredict_Score),2)

ft_dt$Model_Score<- color_bar("lightblue")(ft_dt$AdjPredict_Score)

ft_dt$WAR_Rank <- color_tile("green","orange")(ft_dt$WAR_rank)

ft_dt$Pitching_Rank <- color_tile("green","orange")(ft_dt$AdjPredict_Score_Rank) 


ft_dt$Ranks_Above_ADP <- 
  ifelse(
  ft_dt$Ranks_Above_ADP < 0,
  cell_spec(round(ft_dt$Ranks_Above_ADP,2), color = "red", italic = T),
  cell_spec(round(ft_dt$Ranks_Above_ADP,2), color = "green", italic = T)
)

ft_dt$Positional_Scarcity <- round(ft_dt$Positional_Scarcity, 2)

ft_dt$Pos_Scarcity <- color_tile("red","white")(ft_dt$Positional_Scarcity)

#Name,Team,Position,ADP,WAR, WAR_rank,AdjPredict_Score ,AdjPredict_Score_Rank,Ranks_Above_ADP, Positional_Scarcity

ft_dt2 <- ft_dt[c("Name","Team","Position","ADP", "WAR", "Model_Score", "WAR_Rank","Pitching_Rank","Pos_Scarcity")]


table_export = 
kbl(ft_dt2, escape = F) %>% 
 kable_material(c("striped", "hover","condensed","responsive"),full_width = F,fixed_thead = T) %>%   column_spec(6, width = "3cm")# %>%
 # add_header_above(c(" ", "Scores" = 5, "Ranks" = 3," "))

table_export  


```


### SP
WAR Ranks are for the position, and Pos_Rank is based on Model Score    
```{r, warning =FALSE}

FanGraph_Positions_Pitcher_Total2 =
pitching_Data_NonAdj_Projections_Total %>% 
  mutate(
    Avg_IP = mean(IP),
    AdjPredict_Score_raw = ifelse(is.na(Predict_Score),NA,Predict_Score*(IP/Avg_IP)*1)
  ) %>% select(-Avg_IP) %>%
filter(IP > 19 & PositionSP ==1)


FanGraph_Positions_Pitcher_All_POS2 =
FanGraph_Positions_Pitcher_Total2 %>% 
  mutate(
    max_predscore= max(AdjPredict_Score_raw,na.rm = T),
    AdjPredict_Score = ifelse (is.na(AdjPredict_Score_raw),NA,AdjPredict_Score_raw),
    WAR_rank = order(order(rank(WAR,ties.method = 'average'),decreasing = TRUE)),
    AdjPredict_Score_Rank = order(order(rank(AdjPredict_Score,ties.method = 'average'),decreasing = TRUE))-sum(is.na(AdjPredict_Score)),
        Ranks_Above_ADP = ADP - AdjPredict_Score_Rank
  )  %>%   
  select (Name,Team,Position,ADP,WAR, WAR_rank,AdjPredict_Score ,AdjPredict_Score_Rank,Ranks_Above_ADP, Positional_Scarcity)

ft_dt <- FanGraph_Positions_Pitcher_All_POS2[1:nrow(FanGraph_Positions_Pitcher_All_POS2), 1:ncol(FanGraph_Positions_Pitcher_All_POS2)] %>% 
#%>% 
#filter(AdjPredict_Score_Rank>0)
 arrange((ADP))

ft_dt$ADP <- color_tile("white", "red")(ft_dt$ADP)

ft_dt$WAR <- color_bar("lightblue")(ft_dt$WAR)


ft_dt$AdjPredict_Score<-  round((ft_dt$AdjPredict_Score),2)

ft_dt$Model_Score<- color_bar("lightblue")(ft_dt$AdjPredict_Score)

ft_dt$WAR_Rank <- color_tile("green","orange")(ft_dt$WAR_rank)

ft_dt$Position_Rank <- color_tile("green","orange")(ft_dt$AdjPredict_Score_Rank) 


ft_dt$Ranks_Above_ADP <- 
  ifelse(
  ft_dt$Ranks_Above_ADP < 0,
  cell_spec(round(ft_dt$Ranks_Above_ADP,2), color = "red", italic = T),
  cell_spec(round(ft_dt$Ranks_Above_ADP,2), color = "green", italic = T)
)

ft_dt$Positional_Scarcity <- round(ft_dt$Positional_Scarcity, 2)

ft_dt$Pos_Scarcity <- color_tile("red","white")(ft_dt$Positional_Scarcity)

#Name,Team,Position,ADP,WAR, WAR_rank,AdjPredict_Score ,AdjPredict_Score_Rank,Ranks_Above_ADP, Positional_Scarcity

ft_dt2 <- ft_dt[c("Name","Team","Position","ADP", "WAR", "Model_Score", "WAR_Rank","Position_Rank","Pos_Scarcity")]


table_export = 
kbl(ft_dt2, escape = F) %>% 
 kable_material(c("striped", "hover","condensed","responsive"),full_width = F,fixed_thead = T) %>%   column_spec(6, width = "3cm")# %>%
 # add_header_above(c(" ", "Scores" = 5, "Ranks" = 3," "))

table_export  


```



### RP  
WAR Ranks are for the position, and Pos_Rank is based on Model Score    
```{r,warning=FALSE}

FanGraph_Positions_Pitcher_Total2 =
pitching_Data_NonAdj_Projections_Total %>% 
  mutate(
    Avg_IP = mean(IP),
    AdjPredict_Score_raw = ifelse(is.na(Predict_Score),NA,Predict_Score*(IP/Avg_IP)*1)
  ) %>% select(-Avg_IP) %>%
filter(IP > 19 & PositionRP ==1)


FanGraph_Positions_Pitcher_All_POS2 =
FanGraph_Positions_Pitcher_Total2 %>% 
  mutate(
    max_predscore= max(AdjPredict_Score_raw,na.rm = T),
    AdjPredict_Score = ifelse (is.na(AdjPredict_Score_raw),NA,AdjPredict_Score_raw),
    WAR_rank = order(order(rank(WAR,ties.method = 'average'),decreasing = TRUE)),
    AdjPredict_Score_Rank = order(order(rank(AdjPredict_Score,ties.method = 'average'),decreasing = TRUE))-sum(is.na(AdjPredict_Score)),
        Ranks_Above_ADP = ADP - AdjPredict_Score_Rank
  )  %>%   
  select (Name,Team,Position,ADP,WAR, WAR_rank,AdjPredict_Score ,AdjPredict_Score_Rank,Ranks_Above_ADP, Positional_Scarcity)

ft_dt <- FanGraph_Positions_Pitcher_All_POS2[1:nrow(FanGraph_Positions_Pitcher_All_POS2), 1:ncol(FanGraph_Positions_Pitcher_All_POS2)] %>% 
#%>% 
#filter(AdjPredict_Score_Rank>0)
 arrange((ADP))

ft_dt$ADP <- color_tile("white", "red")(ft_dt$ADP)

ft_dt$WAR <- color_bar("lightblue")(ft_dt$WAR)


ft_dt$AdjPredict_Score<-  round((ft_dt$AdjPredict_Score),2)

ft_dt$Model_Score<- color_bar("lightblue")(ft_dt$AdjPredict_Score)

ft_dt$WAR_Rank <- color_tile("green","orange")(ft_dt$WAR_rank)

ft_dt$Position_Rank <- color_tile("green","orange")(ft_dt$AdjPredict_Score_Rank) 


ft_dt$Ranks_Above_ADP <- 
  ifelse(
  ft_dt$Ranks_Above_ADP < 0,
  cell_spec(round(ft_dt$Ranks_Above_ADP,2), color = "red", italic = T),
  cell_spec(round(ft_dt$Ranks_Above_ADP,2), color = "green", italic = T)
)

ft_dt$Positional_Scarcity <- round(ft_dt$Positional_Scarcity, 2)

ft_dt$Pos_Scarcity <- color_tile("red","white")(ft_dt$Positional_Scarcity)

#Name,Team,Position,ADP,WAR, WAR_rank,AdjPredict_Score ,AdjPredict_Score_Rank,Ranks_Above_ADP, Positional_Scarcity

ft_dt2 <- ft_dt[c("Name","Team","Position","ADP", "WAR", "Model_Score", "WAR_Rank","Position_Rank","Pos_Scarcity")]


table_export = 
kbl(ft_dt2, escape = F) %>% 
 kable_material(c("striped", "hover","condensed","responsive"),full_width = F,fixed_thead = T) %>%   column_spec(6, width = "3cm")# %>%
 # add_header_above(c(" ", "Scores" = 5, "Ranks" = 3," "))

table_export  


```



